---
title: "Grouped plot with weighted density + histogram"
author: "Harald Hofer"
date: "26 2 2020"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Goal of the plot

This plot is suppsed to show the distribution of a KPI for different groups.
The distribution is weighted e.g. using the impressions to take into account the effect of volume.
In addition to that the overall average and average per group is calculated and shown as vertical line.
It also shows the % of volume which are lower than average.


## Background

This kind of plot was created to help investigating a common topic in (digital) media: *Benchmarks*

Benchmarks are (at least implicitly) supposed to give an orientation about how "good" or "bad" a given KPI is. However, misconceptions about what benchmarks (commonly calculated as "overall-average") represent are common. Understanding of the position of the KPI-benchmark compared with the whole distribution of the KPIs is needed.


## The Data

First, get a report or create some data that represents a typical digital ad report which includes Impressions and Clicks for different campaigns, placements, creative sizes and devices for a given time period. For this demo I use the script in my DataGenerators repository.

```{r include = FALSE}
library(tidyverse)

n_campaigns <- 5
n_placements <- 5
n_creativesizes <- 4
n_devices <- 3
n_days <- 60

n_totalrows <- n_campaigns * n_placements * n_creativesizes * n_devices * n_days

# Prepare tibble/dataframe that will be filled with the generated data
tt <- as_tibble(matrix(nrow = n_totalrows, ncol = 8), .name_repair = "minimal")
colnames(tt) <- c("campaign", "placement", "creativesize", "device", "day", "impressions", "clicks", "cost")


# Create campaigns
vec_campaigns <- paste("campaign", LETTERS[1:n_campaigns], sep = "_")


# Create placements for each campaign
list_placements <- vector(mode = "list", length(vec_campaigns))

for (i in 1:length(vec_campaigns)) {
  list_placements[[i]] <- paste(vec_campaigns[i], "placement", round(runif(n_placements, 10000, 99999)), sep = "_")
}

vec_placement <- do.call(c, list_placements)


# Create creatives for each placement
list_creativesizes <- vector(mode = "list", length(vec_placement))

for (i in 1:length(vec_placement)) {
  list_creativesizes[[i]] <- paste(vec_placement[i], "creativeSize", LETTERS[1:n_creativesizes], sep = "_")
}

vec_creativesizes <- do.call(c, list_creativesizes)


# Create devices for each creative
list_devices <- vector(mode = "list", length(vec_creativesizes))

for (i in 1:length(vec_creativesizes)) {
  list_devices[[i]] <- paste(vec_creativesizes[i], "device", LETTERS[1:n_devices], sep = "_")
}

vec_devices <- do.call(c, list_devices)


# Multiply with number of days
list_days <- vector(mode = "list", length(vec_devices))

for (i in 1:length(vec_devices)) {
  list_days[[i]] <- paste(vec_devices[i], "day", 1:n_days, sep = "_")
}

vec_days <- do.call(c, list_days)


# Create impressions
list_impressions <- vector(mode = "list", length(vec_creativesizes))

for (i in 1:length(vec_creativesizes)) {
  list_impressions[[i]] <- round(rbeta(n_days, 1.5, 10000) * 1e7)
}

vec_impressions <- do.call(c, list_impressions)


# Create clicks
list_clicks <- vector(mode = "list", length(vec_creativesizes))

for (i in 1:length(vec_creativesizes)) {
  list_clicks[[i]] <- round(rbeta(n_days, 1.5, 10000) * 1e5)
}

vec_clicks <- do.call(c, list_clicks)



# Fill the tibble with the data
# To fill the categorical dimensions we split the days-vector
tt$campaign <- substr(vec_days, start = 1, stop = 10)
tt$placement <- substr(vec_days, start = 12, stop = 26)
tt$creativesize <- substr(vec_days, start = 28, stop = 41)
tt$device <- substr(vec_days, start = 43, stop = 50)
tt$day <- substr(vec_days, start = 56, stop = 58)
tt$impressions <- vec_impressions
tt$clicks <- vec_clicks
tt$cost <- tt$impressions * rbeta(length(vec_creativesizes), 10, 1e9) * 3e5


# Make adjustments to the clicks per device and creative, so it will better imitate CTR for different devices.
for (i in unique(tt$device)) {
  tt$clicks[tt$device == i] <- tt %>%
    filter(device == i) %>%
    mutate(clicks = round(clicks * runif(1, 0.5, 5))) %>%
    pull(clicks)
}

for (i in unique(tt$creativesize)) {
  tt$clicks[tt$creativesize == i] <- tt %>%
    filter(creativesize == i) %>%
    mutate(clicks = round(clicks * runif(1, 0.5, 5))) %>%
    pull(clicks)
}

tt$clicks[tt$clicks >= tt$impressions] <- 1 # In case this adjustment creates more clicks than there are impressions


# Make adjustments to the costs, so it will be different for different devices and creative sizes
for (i in unique(tt$device)) {
  tt$cost[tt$device == i] <- tt %>%
    filter(device == i) %>%
    mutate(cost = cost * runif(1, 0.5, 5)) %>%
    pull(cost)
}

for (i in unique(tt$creativesize)) {
  tt$cost[tt$creativesize == i] <- tt %>%
    filter(creativesize == i) %>%
    mutate(cost = cost * runif(1, 0.5, 5)) %>%
    pull(cost)
}

```


# The Plot

The plot is built using ggplot2 and includes layers for a density plot and vertical lines with labels that show means of the KPI. This plot is shown in rows for the defines groups.

```{r warnings = FALSE, fig.width = 8, fig.height = 8}
library(tidyverse)

tt_plot <- tt %>%
    group_by(campaign, placement, creativesize, device, day) %>% # This is a preperation in case we want to remove dimensions later
    summarise(sum_clicks = sum(clicks), sum_imp = sum(impressions)) %>% # This is a preperation in case we want to remove dimensions later
    mutate(CTR = sum_clicks / sum_imp) %>%
    as_tibble()

grouped_means <- tt_plot %>% 
    group_by(device) %>%
    summarise(means = sum(sum_clicks) / sum(sum_imp))

grouped_means_pct <- tt_plot %>% 
    group_by(device) %>%
    mutate(means = sum(sum_clicks) / sum(sum_imp)) %>%
    summarise(pct_below = sum(sum_imp[CTR < means], na.rm = TRUE) / sum(sum_imp))
    #summarise(pct_below = sum(CTR < means, na.rm = TRUE) / length(means))

tt_plot %>%
    group_by(device) %>%
    mutate(imp_grp = sum(sum_imp)) %>%
    ggplot(aes(CTR, group = device)) + # , group = device
        geom_vline(aes(xintercept = (sum(sum_clicks) / sum(sum_imp))), color = "grey") +
        geom_label(aes(x = (sum(sum_clicks) / sum(sum_imp)), y = 0.5, label = paste(round((sum(sum_clicks) / sum(sum_imp) * 100), 3),"%", sep = ""), size = 8), color = "grey") +
        # geom_histogram(aes(y = stat(density), fill = device, alpha = 0.7), bins = 30) +
        geom_density(aes(y = stat(density), weight = sum_imp / imp_grp, color = device, fill = device, alpha = 0.7)) + #, adjust = 1
        geom_rug(alpha = 0.1) +
        facet_grid(device ~ .) +
        geom_vline(data = grouped_means, aes(xintercept = means, color = device)) +
        geom_label(data = grouped_means, aes(x = means, y = 0, label = paste(round(means * 100, 3),"%", sep = ""), size = 8, color = device)) +
        geom_label(data = grouped_means_pct, aes(x = 0, y = 0, label = paste(round(pct_below * 100, 1), "%", sep = ""), size = 8, color = device)) +
        xlim(0, quantile(tt_plot$sum_clicks / tt_plot$sum_imp, probs = .75, na.rm = TRUE) + 4 * IQR(tt_plot$sum_clicks / tt_plot$sum_imp, na.rm = TRUE)) +
        theme_minimal()
```

